Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds graphics refs #41

\subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }
This experiment was deployed in a virtual private server with 512MB of memory
runing Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was espected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal aproximations. In order to achieve this, the only change needed 
in the system was to limit the number of chromosmes kept in the pool. 
Previously, every PUT request which included a new chromose increased 
the size of the pool, in a long-running experiment this has to be avoided
in order to not exahust the available memory. In this experiment the size
of a chromosme was 1000 double precision numbers, considereing this
and the memory available in the server, the the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant a random chromosme was pulled out of the pool before inserting a
new one, when the pool was full. For this problem NodEO's \tt{ chromosme-float } 
and \tt{ classic-ea-float} modules were used. The parameters for the ea algorithm 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosme with the server every 100 generations. Again two workers in each window.
As before, a few participation calls were published in social networks.

@@

The results for this new experiment have been extremely interesting
and quite different from the previous one.

>>>>

\subsubsection{Setup of Rastrigin's experiment in NodIO-W$^2$ }
This experiment was deployed in a virtual private server with 512MB of memory
runing Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was espected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal aproximations. In order to achieve this, the only change needed 
in the system was to limit the number of chromosmes kept in the pool. 
Previously, every PUT request which included a new chromose increased 
the size of the pool, in a long-running experiment this has to be avoided
in order to not exahust the available memory. In this experiment the size
of a chromosme was 1000 double precision numbers, considereing this
and the memory available in the server, the the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant a random chromosme was pulled out of the pool before inserting a
new one, when the pool was full. For this problem NodEO's {\tt chromosme-float} 
and {\tt classic-ea-float} modules were used. The parameters for the ea algorithm 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosme with the server every 100 generations. Again two workers in each window.
As before, a few participation calls were published in social networks.

@@

%
\begin{figure}[!htb]
\centering
\includegraphics{rastrigin-fitness.png}
\caption{Fitness of chromosomes sent to the server pool along time in
  the $x$ axis.} 
% Plotted with ../data/plot-rastrigin-fitness.R
\label{fig:puts:rastrigin}
\end{figure}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{rastrigin-IPs.png}
\includegraphics[width=0.49\linewidth]{rastrigin-ips-hour.png}
\caption{Number of unique IPs participating in the Rastrigin
  experiment per minute (left) and per hour (right)} 
% Plotted with ../data/plot-rastrigin-IPs.R
\label{fig:ips:rastrigin}
\end{figure}

The results for this new experiment have been extremely interesting
and quite different from the previous one. The most important
characteristic that set this experiment apart from the previous one is
the fact that it did not find the solution for the time it was
running, instead lowering, little by little, the fitness until it was
disconnected. The fitness of chromosomes sent by clients is shown in
Figure \ref{fig:puts:rastrigin}, that shows two different parts: the
upper part, which represents the fitness of clients that have just
started evolution from a random population, and then the fitness of
those clients that have been already evolving for a while or obtained
a high-fitness individual from the pool and evolved from there, and
which managed to surpass the 9000 value by the end of the
simulation. It incidentally shows also the number of {\em new} nodes
coming: there is a dense {\em cloud} at the beginning, but the new
contributions keep thinning as the evolution progresses.
